---
title: "Forming_Prior_Dist"
format: html
---

```{r}
#| output: false
library(tidyverse)
library(compr)
library(broom)
library(dplyr)
library(readr)
library(deuce)
library(ggplot2)

# Source the functions
source("comp_prior_start.R")
source("bayes_intro.R")
source("wrangle_point_level_data.R")
source("Create_Prior.R")
source("get_probabilities_df.R")
source("get_plot_df.R")
```

We are going to look at the Alcaraz vs Djokovic 2023 Wimbledon Final. Let's include all grass court matches since the start of Wimbledon the year before in 2022, up until the semifinals of the 2023 Wimbledon tournament.

Get probabilities of Alcaraz winning a point while serving and while returning
```{r}
aug_mod_alc <- create_prior(ext = c("atp_matches_2022.csv",
                                 "atp_matches_2023.csv"),
                         tourn_name = "Wimbledon",
                         surf = "Grass",
                         start_date = "2023-01-01",
                         end_date = "2023-07-15",
                         player1 = "Carlos Alcaraz",
                         player2 = "Novak Djokovic",
                         ref_player = "Milos Raonic")
aug_mod_alc

# probability Alcaraz wins a point on serve
expit(0.6194617) # 0.6500961
# probability Alcaraz wins a point on return
expit(-0.6703011) #  0.3384294
# always looking at P1 probability

# Do negation for Djokovic points
# probability Djoko wins a point on serve
expit(0.6703011) # 0.6615706
# probability Djoko wins a point on return
expit(-0.6194617) # 0.3499039
```

Get the point level data for the 2023 Wimbledon final
```{r}
# get ALcaraz vs Djokovic point level data for the 2023 Wimbledon final
# start by getting all point data for 2023 Wimbledon

# NOTE: Player 1 is Alcaraz and Player 2 is Djokovic
both_serving_df <- wrangle_point_level(ext = "2023-wimbledon-points.csv",
                               ID = "2023-wimbledon-1701")

alcaraz_serving <- both_serving_df[[1]]
djokovic_serving <- both_serving_df[[2]]
```

Get a data frame with the probabilities for each player seving
```{r}
combined_prob_df <- get_probabilities_df(p1_serving_df = alcaraz_serving,
                                 p2_serving_df = djokovic_serving,
                                 p1 = "Carlos Alcaraz",
                                 p2 = "Novak Djokovic",
                                 p1_original_prob = 0.6194617,
                                 p1_original_se = 0.1038723,
                                 p2_original_prob = 0.6703011,
                                 p2_original_se = 0.1009560)
```

Now we can plot the probability of Alcaraz winning the match
```{r}
plot_df <- get_plot_df(combined_df = combined_prob_df, 
                       which_player_prob = 1,
                       best_of_3 = FALSE,
                       advantage = FALSE)

set_boundaries <- plot_df |>
  group_by(total_sets) |>
  summarize(xmin = min(pt_number) - 0.5,
            xmax = max(pt_number) + 0.5)

plot_df |> ggplot(aes(x = pt_number, y = probability)) +
  geom_rect(data = set_boundaries, aes(x = NULL, y = NULL, xmin = xmin, xmax = xmax, 
                                       ymin = -Inf, ymax = Inf, fill = total_sets), alpha = 0.2) + 
  geom_line(aes(y = win_prob_px)) +
  labs(x = "Point Number",
       y = "Probability of Winning Match",
       title = "Alcaraz vs Djokovic Wimbledon Final 2023",
       subtitle = "Probability of Alcaraz Winning Match",
       color = "Server") +
  theme_bw()
```